---
title: "Final"
authors: "Nick Yurashku"
date: "2023-05-13"
output: html_document
---

Analysis will start with uploading and cleaning of the data. 


```{r}
#this upload is based off of a pc's directory, be sure to change it to your specific one.

campaigns <- read.csv('C:/Users/nicky/Desktop/Syracuse Grad School/IST 707/Final Project/CandidateSummaryAction.csv')

```


```{r}
#data set needs to be cleaned of data that is not needed:
str(campaigns)
```

Unneeded columns were dropped below

```{r}
campaigns <- subset(campaigns, select = -c(can_off_dis, tot_rec, tot_dis, cas_on_han_beg_of_per,
                                            cas_on_han_clo_of_per, deb_owe_by_com, deb_owe_to_com,
                                            cov_sta_dat, cov_end_dat, can_str1, can_str2, can_cit,
                                            can_zip, off_to_ope_exp, off_to_fun, off_to_leg_acc,
                                            exe_leg_acc_dis, fun_dis, can_loa_rep, oth_loa_rep,
                                            tot_loa_rep, ind_ref, par_com_ref, oth_com_ref,
                                            tot_con_ref, oth_dis, tot_dis))

```


```{r}
str(campaigns)
```
Next, to check for null values:

```{r}
colSums(is.na(campaigns))
```

Votes data is unreliable and will have to be removed as a column. Author of data set said the following: 'the votes column was not part of the metadata. I had to scrape that info manually from CNN's election results page here:

https://www.cnn.com/election/2016/results/house

My apologies for any errors on my part, there is definitely a chance for human error. In which districts have you found discrepancies?'


```{r}
campaigns <- subset(campaigns, select = -c(votes))


```


```{r}
#saving data into a new csv

write.csv(campaigns,'C:/Users/nicky/Desktop/Syracuse Grad School/IST 707/Final Project/campaigns_new.csv')
```



```{r}
#id isn't needed neither
campaigns <- subset(campaigns, select = -c(can_id))
```

```{r}
campaigns <- subset(campaigns, select = -c(can_nam))
```

Need to change data types

```{r}
str(campaigns)
```
before turning everything numeric for money attributes, need to get rid of dollar signs and commas.

```{r}
campaigns$tot_con <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$tot_con)))
campaigns$ind_ite_con <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$ind_ite_con)))
campaigns$ind_uni_con <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$ind_uni_con)))
campaigns$ind_con <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$ind_con)))
campaigns$par_com_con <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$par_com_con)))
campaigns$oth_com_con <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$oth_com_con)))
campaigns$can_con <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$can_con)))
campaigns$tra_fro_oth_aut_com <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$tra_fro_oth_aut_com)))
campaigns$can_loa <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$can_loa)))
campaigns$oth_loa <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$oth_loa)))
campaigns$oth_rec <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$oth_rec)))
campaigns$ope_exp <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$ope_exp)))
campaigns$tra_to_oth_aut_com <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$tra_to_oth_aut_com)))
campaigns$net_con <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$net_con)))
campaigns$net_ope_exp <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$net_ope_exp)))
campaigns$tot_loa <- as.numeric(gsub(",", "", gsub("\\$", "", campaigns$tot_loa)))
```





```{r}
campaigns$can_off <- as.factor(campaigns$can_off)
campaigns$can_off_sta <- as.factor(campaigns$can_off_sta)
campaigns$can_par_aff <- as.factor(campaigns$can_par_aff)
campaigns$can_inc_cha_ope_sea <- as.factor(campaigns$can_inc_cha_ope_sea)
campaigns$can_sta <- as.factor(campaigns$can_sta)

campaigns$winner <- as.factor(campaigns$winner)
```
NA's have been now introduced. So this should be fixed.

```{r}
colSums(is.na(campaigns))
```

All the na's are from null values in the money columns. Turning all of them to zero makes sense.
```{r}
campaigns[is.na(campaigns)] <- 0

```

```{r}
str(campaigns)
```

```{r}
sum(campaigns$tot_con)
```


```{r}
str(campaigns)
```

Candidate state is not needed:

```{r}
campaigns <- subset(campaigns, select = -c(can_sta))
```


```{r}
str(campaigns)
```

```{r}
levels(campaigns$winner)[levels(campaigns$winner) == ""] <- "N"
summary(campaigns$winner)
```

```{r}
levels(campaigns$can_inc_cha_ope_sea)[levels(campaigns$can_inc_cha_ope_sea) == ""] <- "UNKNOWN"
summary(campaigns$can_inc_cha_ope_sea)
```

Printing final cleaned csv

```{r}
write.csv(campaigns,'C:/Users/nicky/Desktop/Syracuse Grad School/IST 707/Final Project/campaigns_cleaned.csv')
```




```{r}
summary(campaigns$can_par_aff)
```

```{r}
ggplot(campaigns, aes(x=can_par_aff)) + geom_bar(aes(y=..count..),fill = 'lightpink',color = 'black',alpha = .8)
```

```{r}
ggplot(campaigns, aes(x=can_par_aff,y=tot_con)) + geom_bar(aes(y=..count..),fill = 'lightpink',color = 'black',alpha = .8)
```


Income sources per party: 
```{r}

df_long2 <- campaigns %>%
  pivot_longer(cols = c(ind_ite_con, ind_uni_con),
               names_to = "contribution_type",
               values_to = "amount")


ggplot(df_long2, aes(fill=contribution_type, y=amount, x=can_par_aff)) + 
  geom_bar(position="stack", stat="identity") +
  labs(x = "Candidate Party Affiliation", 
       y = "Contribution Amount", 
       fill = "Contribution Type", 
       title = "Breakdown of Contributions by Party Affiliation") +
  theme_minimal()


```

Digging deeper, individual contribution break down.

```{r}
df_long <- campaigns %>%
  pivot_longer(cols = c(ind_con, par_com_con, oth_com_con, can_con),
               names_to = "contribution_type",
               values_to = "amount")


ggplot(df_long, aes(fill=contribution_type, y=amount, x=can_par_aff)) + 
  geom_bar(position="stack", stat="identity") +
  labs(x = "Candidate Party Affiliation", 
       y = "Contribution Amount", 
       fill = "Contribution Type", 
       title = "Breakdown of Contributions by Party Affiliation") +
  theme_minimal()
```

Most Expensive State Race

```{r}
ggplot(campaigns, aes(x=can_off_sta,y=tot_con)) + geom_bar(aes(y=..count..),fill = 'magenta',color = 'black',alpha = .8)
#this is too crowded
```

```{r}

# Aggregating the data
state_totals <- campaigns %>%
  group_by(can_off_sta) %>%
  summarise(tot_con_state = sum(tot_con, na.rm = TRUE)) %>%
  arrange(desc(tot_con_state))

# Selecting the top N states
top_states <- state_totals %>%
  top_n(10, wt = tot_con_state)

# Filtering the original data for only these states
campaigns_filtered <- campaigns %>%
  filter(can_off_sta %in% top_states$can_off_sta)

# Create the plot
ggplot(campaigns_filtered, aes(x=can_off_sta, y=tot_con)) + 
  geom_bar(stat="identity", fill = 'magenta', color = 'black', alpha = .8)

```



Party with most wins

```{r}
ggplot(campaigns, aes(x=can_par_aff,y=winner)) + geom_bar(aes(y=..count..),fill = 'darkgoldenrod',color = 'black',alpha = .8)
```





















